---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4

In this assignment we do the following:
- we reproduce the meta-analysis of pitch SD from last semester in a Bayesian framework
- we reproduce the pitch SD in schizophrenia analysis from last semester using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Reproduce the meta-analysis of pitch sd from previous studies of voice in schizophrenia
- the data is available as Assignment4MetaData.xlsx
- Effect size (cohen's d), sd and variance are already calculated (you're welcome!)
- Since we're only interested in getting a meta-analytic effect size, let's take a shortcut and use bromance magic (brms): https://mvuorre.github.io/post/2016/2016-09-29-bayesian-meta-analysis/
```{r}
#libraries
pacman::p_load(readr,groupdata2,ggplot2,tidyverse,data.table, rethinking, metafor,lme4,brms, readxl, rstan)

parallel::detectCores() #4

d_meta <- read_excel("Assignment4MetaData.xlsx")

brm_out <- brm(MeanES|se(SdES)~1+(1|StudyRef), # Outcome as a function of the predictors as in lme4. 
               data=d_meta, # Define the data
               cores=2,
               chain=2,
               #define the family?
               iter = 2000)
summary(brm_out)
plot(brm_out)
#Forrest plot
brmstools::forest(brm_out,
       show_data = TRUE,
       av_name = "Effect size") + geom_vline(aes(xintercept = 0), linetype = "dashed")


#PitchSD~1+ Diagnosis (1+ Diagnosis|P)
#PitchSD~Normal(mu, sigma)
#mu = alpha+ Beta*Diagnosis
#alpha~Normal(0,1) <- leaves it free to vary just as much as orginial data
#beta ~Normal(0,1)
#sigma~cauchy(0,3) - the variance, what is the expected error of our prediction
#log(sigma)~alpha s + beta s * Diagnosis

#mu=alpha [Participant] + beta [Participant]*Diagnosis #each participant will have a different alpha and beta
#alpha[Participant]~ Normal (alpha,1)
#alpha~Normal (0,1)
#you calculate alpha for every participant in the light of alpha of all participant - you pullthem together like that
```


Step 2: Prepare the pitch SD data from last year
- the data is available as Assignment4PitchData.csv (thanks Celine)
- We do not know how to build random effects, yet. So we need to simplify the dataset to avoid interdependence between datapoint: How?
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).
```{r}
#load data
d_pitch <- read_excel("Assignment4PitchDatav2.xlsx")

#simplify the dataset to avoid interdependence between data points
d = d_pitch %>% 
  group_by(ID_unique) %>% 
  mutate(avgPitchSD = mean(PitchSD)) #take the mean across the trials


#let's standardize the data
d <- as.data.frame(d)
d$avgPitchSd.s <- scale(d$avgPitchSD)[,1]
d$PitchMean.s <- scale(d$PitchMean)[,1]

```



Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality

```{r}
#simple model without set priors 
m1 <- brm(avgPitchSd.s ~ diagnosis, 
          data =d,
          family = gaussian(),
          prior = NULL,
          cores = 3, chain = 3, warmup = 2000, iter = 5000)
summary(m1)

temp_df <- posterior_samples(m1)
dens(temp_df$b_diagnosis)
stanplot(m1)
plot(m1)

#define some priors          
m2_formula <- bf(avgPitchSd.s ~ diagnosis + (diagnosis|ID_unique), 
                 sigma ~ diagnosis + (diagnosis|p|ID_unique)) #define formula #?# log sigma?
get_prior(m2_formula, d) #examine priors

prior <- c(prior(normal(0,1), class = Intercept), 
           prior(normal(0,0.5), class = b, coef = diagnosis), 
           prior(normal(0,0.3), class = b, coef = diagnosis, dpar = sigma),
           prior(cauchy(0,1), class = Intercept, dpar = sigma) 
           )

# verify that the priors indeed found their way into Stan's model code
make_stancode(m2_formula,
              data = d, family = gaussian(),
              prior = prior)

#making the model with the priors
m2 <- brm(m2_formula, 
          data = d,
          family = gaussian(),
          prior = prior,
          cores = 1, chain = 1, warmup = 1000, iter = 2000, control = list(adapt_delta = 0.9))
summary(m2)
stanplot(m2)
plot(m2)


#making meta-analytic model
#using a meta analytic prior
prior.m <- c(prior(normal(0,1), class = Intercept), 
           prior(normal(-0.54,0.23), class = b, coef = diagnosis), #skeptical priors
           prior(normal(0,0.3), class = b, coef = diagnosis, dpar = sigma),
           prior(cauchy(0,1), class = Intercept, dpar = sigma) 
           )


m2.m <- brm(m2_formula, 
          data = d,
          family = gaussian(),
          prior = prior.m,
          cores = 1, chain = 1, warmup = 1000, iter = 2000, control = list(adapt_delta = 0.9))

summary(m2.m)
stanplot(m2.m)
plot(m2.m)

```


Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality

Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare their relative distance from truth (WAIC)
- Discuss how they compare and whether any of them is best.
```{r}
#plot priors
x <- seq(-2,2, length=1e5)
y <- dnorm(x, 0, 0.3) #conserv
y.m <- dnorm(x, -0.54, 0.23) #meta
prior_df <- data.frame(x = rep(x,2), y = c(y, y.m), prior = c(rep("conservative", length(y)),
                                                                   rep("meta", length(y.m))
                                                                   ))
ggplot(prior_df, aes(x = x, y = y, color = prior)) + geom_line()+theme_classic()

#plot posterior
post_samples <- c(posterior_samples(m2)$b_diagnosis, posterior_samples(m2.m)$b_diagnosis)
post_df <- data.frame(post_samples = post_samples, model = c(rep("sceptical", 1000),
                                                                   rep("meta", 1000)
                                                                   ))
ggplot(post_df, aes(x = post_samples, color = model)) + geom_density(adjust = 1)+theme_classic()

  #compare models:
waic <- brms::WAIC(m2, m2.m)
waic
pacman::p_load(brms)
weights <- brms::model_weights(m2, m2.m, weights = "waic")
```


Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using WAIC.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?

Optional step 9: Bromance magic.
- explore the bromance code below including random effects (by default with weakly informative priors)
- learn how to change the prior
- explore effects of trial, age, gender, including the appropriate random slopes
- compare the models you created using WAIC and posterior predictive check (pp_check())


```{r}

brm_out <- brm(PitchSD ~ 1 + Diagnosis  +(1|ID_unique/Study), # Outcome as a function of the predictors as in lme4. 
               data=Data, # Define the data
               family=gaussian(), # Define the family. 
               iter = 5000, warmup = 2000, cores = 4)
summary(brm_out1)
plot(brm_out1)

```

